---
title: "Exploring bupaR"
output: html_document
theme: cosmo
---

# Load Libraries and Create Event Log Object from csv File

```{r setup, include=TRUE}
# Load libs
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(lubridate))

# Import event log form csv
event_log_df <-
  read.csv(here::here("Data", "event_log.csv"),
           stringsAsFactors = FALSE,
           sep = ";")

# Select life prod, add activity_instance_id, add lifecycle_id
event_log_df_life <- event_log_df %>%
  filter(PRODLINE == "life") %>%
  select(-PRODLINE) %>%
  mutate(ACTIVITY_INST_ID = as.numeric(row.names(.))) %>%
  gather(LIFECYCLE_ID,
         TIMESTAMP,
         -PROPOSALID,-USERID,
         -ACTIVITY_INST_ID,
         -EVENT) %>%
  mutate(TIMESTAMP = ymd_hms(TIMESTAMP)) %>%
  arrange(PROPOSALID, TIMESTAMP) %>%
  mutate(LIFECYCLE_ID = case_when(.$LIFECYCLE_ID == 'STARTTIMESTAMP' ~ 'START',
                                  TRUE ~ 'END'))

# Transform df into bupaR eventlog
suppressPackageStartupMessages(library(bupaR))
suppressPackageStartupMessages(library(processmapR))

event_log_life <- event_log_df_life %>%
  eventlog(
    case_id = "PROPOSALID",
    activity_id = "EVENT",
    activity_instance_id = "ACTIVITY_INST_ID",
    lifecycle_id = "LIFECYCLE_ID",
    timestamp = "TIMESTAMP",
    resource_id = "USERID"
  )
```

# Exploratory Data Analysis of the Event Log (EDA)

## Explore Event Log Object Mappings
```{r}
event_log_life %>% mapping
```

## Summary of Event Log Object
```{r}
event_log_life %>% summary
```


## Graph of Processing Times per Event

```{r}
event_log_life %>% 
    processing_time("activity") %>%
    plot
```


## Graph of Throughput Times

This can be measured at various levels, by default here on the whole log.

```{r}
event_log_life %>% 
     throughput_time("log")%>%
    plot
```

## Measures of Structuredness

### Activity Presence
From docs: "Activity presence shows in what percentage of cases an activity is present."

```{r}
event_log_life %>%  activity_presence() %>%
    plot
```

### Activity Frequency

```{r}
event_log_life %>%
    activity_frequency("activity")
```

### Trace Coverage
From docs: "The trace coverage metric shows the relationship between the number of different activity sequences (i.e. traces) and the number of cases they cover."

```{r}
event_log_life %>%
    trace_coverage("trace") %>%
    plot()
```